Fractional Types

### 2016-12-18 typespp.tex
clarified example

The first step is to enrich the types to pointed types $\mathbb{T_\bullet}$
where each type $\mathbb{T}$ refers to a particular value $v$ in
$\mathbb{T}$. We recall that, as all fractional types are inhabited (see remark
in Sec.~\ref{sec:symmetryg}), and thus there is no loss in generality in going
to pointed types. We then allow universal quantification over indices $k$
associated to values, over a collection of types.  In particular, the type
$\forall k.[\order{c} \circledast \iorder{c}, \langle c^k, 1/c^k \rangle]$
imposes a constraint on the two values associated with the types $\order{c}$ and
$\iorder{c}$; they must be ``synchronized'' by sharing the same $k$ but are
otherwise arbitrary. Once introduce a universal quantifier in one direction of
execution, we are led to also include an existential quantifier for the reverse
execution. Intuitively, a forward evaluation step that is at liberty to choose a
$k$ will, when reversed encounter a particular, i.e., existentially quantified,
$k$. The syntax is therefore as follows.

\begin{definition}[Dependent $\Pi^/$]
\label{def:pifracd}
The syntax of dependent $\Pi^/$ differs from Def.~\ref{def:pifrac} as follows:
\[\begin{array}{lrcl}
(\textrm{Fractional Types}) & \mathbb{T_\bullet} &::=&
  [\order{c},c^k] \alt
  [\iorder{c},1/c^k] \alt
  [\mathbb{T}_1 \circledast \mathbb{T}_2,\langle v_1,v_2 \rangle] \alt
  \forall k.\mathbb{T_\bullet} \alt \exists k.\mathbb{T_\bullet} \\
 (\textrm{/-combinators}) & \rho_\bullet &:&
  \mathbb{T}_{\bullet 1} \leftrightblackspoon \mathbb{T}_{\bullet 2} ~::=
  [\textit{see Fig.~\ref{pifracd:comb}}]
\end{array}\]
\end{definition}

Here is how the circuit in Fig.~\ref{fig:zigzag} evaluates in the forward and
backwards directions:
\[\begin{array}{rcl}
[\order{c},c^i]
&\stackrel{\unitetl/}{\longmapsto}&
  [\order{id}\circledast\order{c}, \langle \idiso,c^i \rangle] \\
&\stackrel{\eta+\circledast\idiso/}{\longmapsto}&
  \forall{k}.[(\order{c}\circledast\iorder{c})\circledast\order{c},
  \langle \langle c^k,1/c^k \rangle,c^i \rangle] \\
&\stackrel{\assocrt/}{\longmapsto}&
  \forall{k}.[\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^k, \langle 1/c^k,c^i \rangle\rangle] \\
%% &\stackrel{\synch_i}{\longmapsto}&
&k\mapsto i&
  [\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^i, \langle 1/c^i,c^i \rangle\rangle] \\
&\exists&
  \exists{k}.[\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^i, \langle 1/c^k,c^k \rangle\rangle] \\
&\stackrel{\idiso/\circledast\epsilon-}{\longmapsto}&
  [\order{c}\circledast\order{\idiso},
  \langle c^i, \idiso \rangle] \\
&\stackrel{\unitetr/}{\longmapsto}&
  [\order{c}, c^i]
\end{array}\]

@@

Need to explain that a $\forall$ in the forward direction is a choice and
becomes $\exists$ in the backwards direction.

>>>>

The first step is to enrich the types to pointed types $\mathbb{T_\bullet}$
where each type $\mathbb{T}$ refers to a particular value $v$ in
$\mathbb{T}$. We recall that, as all fractional types are inhabited (see remark
in Sec.~\ref{sec:symmetryg}), and thus there is no loss in generality in going
to pointed types. We then allow universal quantification over indices $k$
associated to values, over a collection of types.  In particular, the type
$\forall k.[\order{c} \circledast \iorder{c}, \langle c^k, 1/c^k \rangle]$
imposes a constraint on the two values associated with the types $\order{c}$ and
$\iorder{c}$; they must be ``synchronized'' by sharing the same~$k$ but are
otherwise arbitrary. Once we introduce a universal quantifier in one direction of
execution, we are led to also include an existential quantifier for the reverse
execution. Intuitively, a forward evaluation step that is at liberty to choose a
$k$ will, when reversed encounter a particular, i.e., existentially quantified,
$k$. The syntax is therefore as follows.

\begin{definition}[Dependent $\Pi^/$]
\label{def:pifracd}
The syntax of dependent $\Pi^/$ differs from Def.~\ref{def:pifrac} as follows:
\[\begin{array}{lrcl}
(\textrm{Fractional Types}) & \mathbb{T_\bullet} &::=&
  [\order{c},c^k] \alt
  [\iorder{c},1/c^k] \alt
  [\mathbb{T}_1 \circledast \mathbb{T}_2,\langle v_1,v_2 \rangle] \alt
  \forall k.\mathbb{T_\bullet} \alt \exists k.\mathbb{T_\bullet} \\
 (\textrm{/-combinators}) & \rho_\bullet &:&
  \mathbb{T_1}_{\bullet} \leftrightblackspoon \mathbb{T_2}_{\bullet} ~::=
  [\textit{see Fig.~\ref{pifracd:comb}}]
\end{array}\]
\end{definition}

Here is how the circuit in Fig.~\ref{fig:zigzag} evaluates in the forward and
backwards directions where we diambiguate the uses of $\forall$ and $\exists$
for clarity.

\[\begin{array}{rcl}
[\order{c},c^i]
&\stackrel{\unitetl/}{\longmapsto}&
  [\order{id}\circledast\order{c}, \langle \idiso,c^i \rangle] \\
&\stackrel{\eta+\circledast\idiso/}{\longmapsto}&
  \forall{k}.[(\order{c}\circledast\iorder{c})\circledast\order{c},
  \langle \langle c^k,1/c^k \rangle,c^i \rangle] \\
&\stackrel{\assocrt/}{\longmapsto}&
  \forall{k}.[\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^k, \langle 1/c^k,c^i \rangle\rangle] \\
%% &\stackrel{\synch_i}{\longmapsto}&
&k\mapsto i&
  [\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^i, \langle 1/c^i,c^i \rangle\rangle] \\
&\exists&
  \exists{k}.[\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^i, \langle 1/c^k,c^k \rangle\rangle] \\
&\stackrel{\idiso/\circledast\epsilon-}{\longmapsto}&
  [\order{c}\circledast\order{\idiso},
  \langle c^i, \idiso \rangle] \\
&\stackrel{\unitetr/}{\longmapsto}&
  [\order{c}, c^i]
\end{array}\]

@@
